Open-source landscape for 3D CSEM modelling

### 2020-07-06 manuscript/Manuscript.tex
Incorporating LH's changes; start to rework.

% Remove once done
\usepackage[textsize=footnotesize, textwidth=2.8cm]{todonotes}
% \usepackage[textsize=footnotesize, textwidth=2.8cm, disable]{todonotes}

% Additional packages to geophysics.cls
% \usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

\usepackage{amssymb, amsmath, amsfonts}
\usepackage{upquote}
\usepackage[strings]{underscore}
\usepackage{siunitx}                % SI conform commands (eg \num)
\usepackage{tabularx}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{flafter}

\usepackage[pdftex, final]{hyperref}
% \usepackage[pdftex, hidelinks]{hyperref}
\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}

% Figures
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\renewcommand{\figdir}{./figures}
\ifthenelse{\boolean{@twoc}}{%
  \newcommand{\cwidth}{250pt}}{%
  \newcommand{\cwidth}{240pt}}

\ifthenelse{\boolean{@twoc}}{%
  \newcommand{\acwidth}{250pt}%
  \newcommand{\tcwidth}{250pt}}{%
  \newcommand{\acwidth}{240pt}%
  \newcommand{\tcwidth}{312pt}}

% Own commands
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\emg}[2]{\texttt{emg#1#2}\xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\newcommand{\simpeg}{\texttt{SimPEG}\xspace}
\newcommand{\custem}{\texttt{custEM}\xspace}
\newcommand{\petgem}{\texttt{PETGEM}\xspace}

% Remove once done
\newcommand{\mycom}[2][]{
  \todo[color=yellow]{\textbf{\uppercase{[#1]}}:~#2}}
\newcommand{\imycom}[2][]{
  \todo[inline, color=yellow]{\textbf{\uppercase{[#1]}}:~#2}}
\newcommand{\itodo}[1]{\todo[inline, color=cyan]{\sffamily #1}}
\newcommand{\dul}[1]{\underline{\underline{#1}}}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\rmk}[1]{{\color{red}{#1}}}
\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}

@@

\address{
\footnotemark[1]TU Delft,
Building 23,
Stevinweg 1 / PO-box 5048,
2628 CN Delft;\\
\footnotemark[2]TODO\\
\footnotemark[3]TODO\\
\footnotemark[4]University of California Berkeley, Department of Statistics\\
E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org};\\
\textbf{Keywords}: CSEM, Open-Source, 3D modelling, FD, FE.
}

@@

%%fakesection ===    ABSTRACT    ===
\begin{abstract}
%
\emph{I will draft a first abstract after the manuscript has been reviewed by
everyone once.}
%
% % Principal objectives and scope of the work
% % Methodology
% % Results
% % Conclusions
%
\end{abstract}

\itodo{
  For comments in the margin, use:\\[.2cm]
  \qquad\texttt{\textbackslash mycom[YOURINITIALS]\{comment\}}\\[.2cm]
  and for inline comments, use:\\[.2cm]
  \qquad\texttt{\textbackslash imycom[YOURINITIALS]\{comment\}}\\[.5cm]
  If you want to compile it without comments, comment the line\\[.2cm]
  \qquad\texttt{\textbackslash usepackage[textsize=footnotesize, textwidth=2.8cm]\{todonotes\}}\\[.2cm]
  and uncomment the line\\[.2cm]
  \qquad\texttt{\% \textbackslash usepackage[textsize=footnotesize, textwidth=2.8cm, disable]\{todonotes\}}
}

\itodo{
We will have to decide on a Journal. Current suggestions:\\
- GMD: Geoscientific Model Development\\
- SIG: Surveys in Geophysics\\
- SE: Solid Earth\\
- GJI: Geophysical Journal International\\
- GP: Geophysical Prospecting\\
- GEO: Geophysics\\
- CAG: Computers and Geosciences\\
\imycom[DW]{
I think \emph{Geoscientific Model Development} and \emph{Surveys in Geophysics}
were the favourite ones when we discussed it ones. I personally would be very
interested in GMD. What do you think? But also interesting would be the GJI,
because the MT comparison was published there, so it would be in the same
place. I just had a very bad experience with GP, so I am not inclined on that
at all.}
\imycom[RR]{GMD looks very interesting, my only concern is that it might not be
in the scope of EM modelers, but we can probably reach modeleres from other
communities as well.}
\imycom[OCR]{I also find GMD very interesting, in fact it was my first option to
publish the first petgem version (2 years ago). The answer was: "manuscript
outside of journal's scope". In this sense I agree with Raphael. My vote is for
SIG}
}
\imycom[LH]{SIG does look like a good option. Given the spirit of the paper, it
would be nice to publish it open-access. Does anyone have a sense of how much that
costs with SIG? I could likely get partial funding towards this. Solid Earth
might be a nice option in that case because it is open-access by default. }

@@

\imycom[RR]{We need to take care to not mix British English with AE, I'm very
  used to AE, but Dieter started with (at least what I can identify) mainly BE,
  so I tried to follow (mainly s instead of z and ou instead of o in some
  words). I hope there are not too many inconsistencies, as probably Lindsey
  has to fix all these mistakes in the end:-)
\imycom[DW]{Be aware of Canadian English and Australian English :-D -- You are
  of course right. I am usually mostly BE, having done my PhD in Scotland.
  However, I don't think it matters at the moment, because it will depend to
  what journal we submit. With an appropriate spellchecker that should be a
  quick thing then.}}
\imycom[LH]{It's true! Canadian English ends up being an interesting hybrid(!)}

\imycom[OCR]{In general, I believe that the content is adequate and the ideas
are clear and in accordance with the objective that we have initially set.
However, I want to highlight the task of being careful when justifying the
absence of the SimPEG results for the Marlim model (I have read it in Slack,
although it may not be definitive). I believe that this may or may not be a
weak point depending on how we explain it. I know that perhaps it is not
serious, I only make this comment so that the speech justifies the facts. }

Controlled-source electromagnetic (CSEM) measurements are a frequently applied
method in various geophysical exploration fields, such as geothermal,
groundwater, oil and gas, mining, civil engineering, or geo-hazards. Modelling
these electromagnetic fields is therefore of great interest to design survey
layouts, to understand the measured data, and for inversion purposes.
Publications regarding 3D modelling in electromagnetic methods started to
appear as early as the 1970's and 1980's. These early publications were
integral equation (IE) methods, having an anomaly embedded within a layered
medium, mostly for loop-loop type transient EM measurements
\citep{GJI.74.Raiche, GEO.75.Hohmann, GJI.82.Das, GEO.86.Newman} and
magnetotelluric (MT) measurements \citep{GEO.84.Wannamaker}. \cite{Ward1988}
assemble many of these solutions in ``Electromagnetic Theory for Geophysical
Applications,'' which is widely viewed as an authoritative publication on
electromagnetic geophysics.

@@

\imycom[LH]{I added a couple citations here - I think the Haber 2001 is important to include in the FV list (if we want to remove the 2007 one to reduce duplicating authors, I think that is okay as we cite it later)}

The continuous improvement of computing power and the CSEM boom in the early
2000's in the hydrocarbon industry led to a wealth of developed numerical
solutions and according publications. The most common applied methods to solve
Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan,
GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and
different variations of the differential equation (DE) method, such as finite
differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin,
GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer,
GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev,
SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, Haber2001,PIER.01.Clemens,
GP.06.Mulder, ECP.07.Haber, GEO.14.Jahandari}. There are also many different
types of discretisation, where the most common ones are regular grids
(cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a
Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids
\citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or
OcTree meshes \citep{ECP.07.Haber}. Very well written overviews about the
different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and
\cite{SG.10.Borner}.

@@

\imycom[LH]{Here, I think that there
is a point to make about duplication of efforts vs. new methodologies. Often
the development of a new methodology requires that a new implementation be
built from scratch. As a field though, we have matured to a point where we
know what the methods / approaches are, and much of the next stage of
advancements will be more incremental. (We argued this in Oldenburg et al,
2019 - so to self-cite ;) (doi: https://doi.org/10.1080/08123985.2019.1580118,
arxiv: https://arxiv.org/pdf/1902.08245.pdf). See what you think of the
following framing for the next paragraph - we can discuss:

Many of the advancements made in the EM modelling community over the past
decades have required that authors develop new implementation from scratch.
These codes often provided the research group or company with a competitive
advantage for a time, and thus the source codes were often kept internal. In
some cases, executables have been made available for academic purposes upon
request or to sponsoring consortium members. As the field continues to mature,
advancements become more incremental, and particularly in an applied field,
many advancements are driven by new use-cases and applications that were not
considered by the original authors. In his review on EM modelling and
inversion \cite{SG.05.Avdeev} concludes with the
following statement: «\emph{The most important challenge that faces the EM
community today is to convince software developers to put their 3-D EM forward
and inverse solutions into the public domain, at least after some time. This
would have a strong impact on the whole subject and the developers would
benefit from feedback regarding the real needs of the end-users.}» Further,
\cite{Oldenburg2019} argue that an open-source paradigm has the potential to
accelerate multidisciplinary research by encouraging the development of
modular, interoperable tools that are supported by a community of researchers.
}

Probably the main reason for the numerous amount of code developments was
outlined by \cite{SG.05.Avdeev}, who finished his review with the
following statement: «\emph{The most important challenge that faces the EM
community today is to convince software developers to put their 3-D EM forward
and inverse solutions into the public domain, at least after some time. This
would have a strong impact on the whole subject and the developers would
benefit from feedback regarding the real needs of the end-users.}» Some major
exploration service companies, as well as certain research consortia and
research groups, have their own 3D CSEM codes, but how about codes that are
open-source? Academic codes were traditionally often available upon request by
authors. Codes distributed that way are often provided with the request that users not
share the code, and they often come without a license attached which means that
they are, legally, not open-source. Also, at times there can be significant
hurdles to install a code.

\imycom[LH]{Here, I think it is worth discussing the distinction between
open-source: source code available + openly licensed (with an Open Source
Initiative (OSI, https://opensource.org) approved license), open-development:
code is versioned, development and issue tracking happens publicly (e.g. on
GitHub or similar), and open communities: actively soliciting contributions
and pull requests which change the code. and then the push towards open
development / open communities motivates best practices including continuous
integration, documentation, etc. See what you think of my suggested adaptation
below. There might be a few citations we can add for an open model of development / open community project. I will do some digging. Again, happy to discuss.

In many domains of science, it is becoming more common for researchers to
release source-code with an open license that allows use and modification (
e.g. see the Open Source Initiative approved licenses:
\href{https://opensource.org/licenses}{https://opensource.org/licenses}). This
is an important step for improving reproducibility of research: «\emph{to
provide the means by which the reader can verify the validity of the results
and make use of them in further research}» \citep{GEO.17.Broggini}. Going a
step beyond releasing open-source software, many groups adopt an open model of
development where code is hosted and versioned in an online repository,
all changes are public, and users can engage by submitting and track issues.
Community oriented projects further engagement by encouraging pull-requests,
which are suggested changes to the code. Successful, well-maintained projects
often have unit-testing and continuous integration that runs those tests with
any changes to the code. Additionally, documentation that includes examples
and tutorials is an important component for on-boarding new users and
contributors. As a result, in many areas of the geosciences, we are seeing a
shift away from a one-way distribution of code towards building global
communities around open projects. Related projects within the
realm of geophysical exploration are e.g., \texttt{pyGIMLi}
\citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}
}

The meaning of the term open-source evolved rapidly in the last two decades.
Today, open-source generally not only means that a code is available (with a
proper license), but also that the code is hosted online, is under version
control, has the possibility to file issues and make pull requests to fix bugs.
These are important requirements for reproducible research «\emph{to provide
the means by which the reader can verify the validity of the results and make
use of them in further research}» \citep{GEO.17.Broggini}. Well maintained
codes often have continuous integration that includes testing of the code base
and online hosted documentation, often including examples and tutorials. As
such, the term has shifted from purely open-source codes to developments in the
open with increasingly building a global community. Related projects within the
realm of geophysical exploration are e.g., \texttt{pyGIMLi}
\citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}.
\imycom[DW]{Others coming to your mind? An interesting looking one is
  \href{https://github.com/fancompute/ceviche}{github.com/fancompute/ceviche},
  but they are at very high frequencies on nm-distances.
\imycom[RR]{no thoughts about geophysics, but there are many projects for
  related topics such as gmt, gis, mesh-gen}
\imycom[OCR]{No idea from my side}
\imycom[LH]{Could we cite: https://github.com/softwareunderground/awesome-open-geoscience ?. Maybe we could ask them to put it on zenodo so it has a doi?? Also Madagascar in seismic?}
}

In this paper, we discuss the topic of open-source developments with respect to 3D
CSEM modeling. Even though it comprises still the minority of developments, the
landscape in this field of research changed in the last five years. Previously,
there were only closed-source codes owned by companies or  consortia and
individual projects, often without proper documentation, that could be
requested from the authors. We introduce and compare four recent open-source
projects with focus on 3D CSEM modeling. All presented codes are in the Python
ecosystem and use either the FV method on structured grids or the FE method on
unstructured tetrahedral meshes.

@@

We simulate EM fields for a layered background model with vertical transverse
isotropy, containing three blocks, as well as for the complex marine Marlim R3D model.
These two model designs as well as the corresponding results from four
different codes provide a benchmark for new codes to be compared to and
validated with. In the following section, we introduce the codes under
consideration. Afterwards, we present the considered benchmark cases in detail
and present the modeling results of our four codes in terms of accuracy and
computational performance. Beyond that, we extensively discuss important points
that control the performance and suitability of our FV and FE codes, including
mainly considerations about the mesh design and the choice of solvers. We
conclude with a summary of our results and a motivation for the EM community to
not only continue to extend the landscape of open-source codes but to also
create a landscape of open-source benchmark models.

\imycom[RR]{Regarding
  \todo[inline, color=white]{
  "It is worth mentioning that there are many more 3D EM
  codes developed in other fields than geophysics such as communication systems
  (antennas, radar, satellites), medical imaging, or civil engineering. There
  are various open-source codes in these fields too, e.g., Elmer
  (\href{http://www.csc.fi/elmer}{csc.fi/elmer}, couldn't
  find a publication, just a poster).
  \todo[inline, color=gray]{(\textbf{[DW]}: Add 2-3 proper references)}
  While these codes could potentially be used for the same goals as presented
  here we restrict our review to codes purpose-built and ready without further
  adjustments for geophysical applications."
}
  I thought a lot about this but I just think the previous paragraph is not
  required in our introduction and haven't find a location where it fits
  considering different logical structures. I think it is also not important
  for this paper to tell "that there are more codes in EM than only
  geophysics", that's quite obvious. Maybe delete or shorten to 1 sentences to
  add Elmer and place this sentence somewhere else? I think, an appropriate
  list of all non-geophysical open-source EM codes should be very long anyway,
  we just don't know them or at least I'm not familiar with them.}

\imycom[LH]{Just a thought, but right now we don't state Maxwell's equations anywhere! That seems like something we should do :) That also provides an easy space for talking about Fourier Transform conventions and such too.

% Following the conventions of Ward and Hohmann \cite{Ward1988}, the quasi-static Maxwell's equations are

% \begin{equation}
% \begin{split}
%     \nabla \times \vec{E} &= -i \omega\vec{B}  \\
%     \nabla \times \vec{H} &= \vec{J}
% \end{split}
% \label{eq:maxwell_freq_quasi}
% \end{equation}
}

\clearpage  % TODO: Just for now while drafting, remove later!
\section{Codes}

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg
\citep{CAG.15.Cockett, Heagy2017}. All four codes have their user-facing routines written
in Python; all of them make heavy use of \texttt{NumPy}
\citep{CSE.11.VanDerWalt} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four
of them are “modern” open-source projects, meaning that they come with both, an
open-source license and an online-hosted version-control system with tracking
possibilities (raising issues, filing pull requests). All developments comprise
an extensive online documentation with many examples and have continuous
integration to some degree. Newer package-management systems such as Conda,
Docker, or pip significantly supported us to make our codes easily available
and installable without any user-side compilations. Each one can be downloaded
and installed with a single command.

In the following, we provide a quick overview of the codes. It is, however, beyond the scope
of this article to go into every detail of the different modellers, and we
refer to their documentations for more details. An overview comparison of the
different codes is given in Table~\ref{tbl:codecomparison}. All codes have in
common that they solve the\mycom[DW]{are we all using the weak form?}\mycom[LH]{SimPEG uses weak form} weak
formulation of Maxwell's equation in its differential form (DE) under the
quasistatic or\mycom[DW]{are we all using the diff. approx?}\mycom[OCR]{petgem
consider the weak work of maxwell equations in their diffusive form
(quasi-static approximation)}\mycom[LH]{SimPEG uses quasistatic approx.} diffusive approximation, hence neglecting
displacement currents. The machines on which the different codes were run are
listed in Table~\ref{tbl:machines}\mycom[LH]{SimPEG also interfaces to the scipy solvers (but in practice those are use much less than Pardiso / mumps)}, together with the responsible operator.

@@

\itodo{
Each code should be briefly summarized by three points:\\
- Brief intro\\
- Main selling point\\
- Planned features\\
I think in a first round \textbf{everyone should just write what they want to
write and think is important for their code}. In a second round we should then
homogenize it, such that all codes have more or less the same length/weight and
also similar topic/layout.
}

\clearpage  % TODO: Just for now while drafting, remove later!
\subsection{custEM}

The customisable electromagnetic modeling Python toolbox custEM was developed
for simulating arbitrary complex 3D CSEM geometries with focus on semi-airborne
setups, but it supports also land-based, airborne, coastal  or marine
environments. Multiple electric or magnetic field or potential finite-element
approaches were implemented as total or secondary field formulations. The
finite-element kernel, including higher order basis functions and
parallelisation, relies on the FEniCS project \citep{B.SPR.12.Logg,
B.SPR.16.Langtangen}. The resulting systems of linear equations are solved with
MUMPS \citep{SIAM.01.Amestoy}, which is a very robust but memory consuming
choice. Primary field solutions are supplied by the COMET \citep{GEO.20.Skibbe}
package.

The toolbox considers generally anisotropic petrophysical properties. Even
though changes of the conductivity is mainly of interest for CSEM modeling, the
electric permittivity and magnetic permeability can be taken into account using
the preferred electric field approach on Nédélec elements. Recently, induced
polarization parameters in frequency-domain and three methods for simulating
time-domain responses were added to custEM. The provided meshing tools based on
TetGen \citep{TOM.15.Si} and functionalities of pyGIMLi \citep{CAG.17.Rucker}
facilitate the generation of tetrahedral meshes including layered-earth
geometries with topography or bathymetry and anomalous bodies which are allowed
to be connected or to reach the surface.

The custEM toolbox is under steady development and maintenance. Future
enhancements comprise adding iterative solution techniques, provided by FEniCS,
as well as further functionalities for automatically incorporating intersecting
bodies such as folds or faults in meshes. Furthermore, current work is on using
custEM for inverse modeling applications.

@@

Based on the current state of the \petgem project, there are many possibilities
to improve or to add additional features. For instance, in the short-term, more
modeling routines and support for multi-source set-ups will be implemented. For
the long-term, the obvious next step is to move to 3D CSEM data inversion.

@@

\simpeg is a modular toolbox for simulations and gradient based inversions in
geophysics. Current functionality includes modelling and inversion
capabilities for gravity, magnetics, direct current resistivity, induced
polarization, electromagnetics (time, frequency, as well as natural sources,
e.g. magnetotellurics), and fluid flow (Richards equation). It is a community
driven project that aims to support researchers by providing a flexible,
extensible framework and to facilitate integration of data sets from a variety
of geophysical methods, including joint inversions \citep{Astic2020}, by
providing a common interface to each method.

>>>>

% TODO's
\usepackage[color=cyan, textsize=footnotesize, textwidth=2.8cm]{todonotes}

% Additional packages
\usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{siunitx}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{flafter}
\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}

% Figures
\renewcommand{\figdir}{./figures}

% Own commands
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\emg}[2]{\texttt{emg#1#2}\xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\newcommand{\simpeg}{\texttt{SimPEG}\xspace}
\newcommand{\discretize}{\texttt{discretize}\xspace}
\newcommand{\custem}{\texttt{custEM}\xspace}
\newcommand{\petgem}{\texttt{PETGEM}\xspace}
\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}

% TODO's
% For comments in the margin, use:   \mycom[YOURINITIALS]{comment}
% and for inline comments, use:      \imycom[YOURINITIALS]{comment}
\newcommand{\mycom}[2][]{\todo{\textbf{\uppercase{[#1]}}:~#2}}
\newcommand{\imycom}[2][]{\todo[inline]{\textbf{\uppercase{[#1]}}:~#2}}
\newcommand{\itodo}[1]{\todo[inline]{\sffamily #1}}

@@

\address{
\footnotemark[1]%
TU Delft,
Building 23,
Stevinweg 1 / PO-box 5048,
2628 CN Delft (NL);\\
\footnotemark[2]%
Leibniz Institut für Angewandte Geophysik,
Stilleweg 2,
30655 Hannover (DE);\\
\footnotemark[3]%
Barcelona Supercomputing Center (BSC),
Nexus II Building c/Jordi Girona, 29,
08034 Barcelona (ES);\\
\footnotemark[4]%
University of California Berkeley,
Department of Statistics (USA).\\
E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org};\\
\textbf{Keywords}: CSEM, Open-Source, 3D modelling, FV, FE.
}

@@

%%fakesection ===    ABSTRACT    ===
\begin{abstract}
%
Large-scale modelling of three-dimensional (3D) controlled-source electromagnetic (CSEM) used to be only possible for large companies or research consortia. However, this changed over the last few years, and today there exists a selection of different open-source codes which are up to the task.
%
Four different open-source 3D CSEM codes are considered in this work, where in all cases at least the user-facing functionality is written in Python. Two codes, \simpeg and \emg3d, use finite volumes, the other two codes, \custem and \petgem, use finite elements. Three different mesh-types are used, simple tensor meshes, unstructured tetrahedral meshes, and octree meshes. Two models are used to compare the codes, a layered background containing three resistive blocks, and Marlim R3D, a complex, real-world, open-source resistivity data.
%
The comparison shows that all four codes are able to compute the model to sufficient precision. The biggest difference between the codes are actually the used mesh types. The required runtime and memory varies greatly between the codes. However, these requirements depend mainly on the underlying solver, and not on the actual CSEM codes.
%
The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community, and we invite and encourage the community to make more code, modelling scripts, and results publicly available. For a start, all the scripts to reproduce our shown results are available.
%
\end{abstract}

@@

Controlled-source electromagnetic (CSEM) measurements are a frequently applied method in various geophysical exploration fields, such as geothermal, groundwater, oil and gas, mining, civil engineering, or geo-hazards. Modelling these electromagnetic fields is therefore of great interest to design survey layouts, to understand the measured data, and for inversion purposes. Publications regarding 3D modelling in electromagnetic methods started to appear as early as the 1970's and 1980's. These early publications were integral equation (IE) methods, having an anomaly embedded within a layered medium, mostly for loop-loop type transient EM measurements \citep{GJI.74.Raiche, GEO.75.Hohmann, GJI.82.Das, GEO.86.Newman} and magnetotelluric (MT) measurements \citep{GEO.84.Wannamaker}. \cite{B.SEG.88.Ward} assemble many of these solutions in \emph{Electromagnetic Theory for Geophysical Applications}, which is widely viewed as an authoritative publication on electromagnetic geophysics.

@@

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most common applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretisation, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or OcTree meshes \citep{ECP.07.Haber}. Very well written overviews about the different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and \cite{SG.10.Borner}.

@@

Many of the advancements made in the EM modelling community over the past decades have required that authors develop new implementation from scratch. These codes often provided the research group or company with a competitive advantage for a time, and thus the source codes were often kept internal. In some cases, executables have been made available for academic purposes upon request or to sponsoring consortium members. As the field continues to mature, advancements become more incremental, and particularly in an applied field, many advancements are driven by new use-cases and applications that were not considered by the original authors. In his review on EM modelling and inversion \cite{SG.05.Avdeev} concludes with the following statement: «\emph{The most important challenge that faces the EM community today is to convince software developers to put their 3-D EM forward and inverse solutions into the public domain, at least after some time. This would have a strong impact on the whole subject and the developers would benefit from feedback regarding the real needs of the end-users.}» Further, \cite{EXG.19.Oldenburg} argue that an open-source paradigm has the potential to accelerate multidisciplinary research by encouraging the development of modular, interoperable tools that are supported by a community of researchers.

In many domains of science, it is becoming more common for researchers to release source-code with an open license that allows use and modification (e.g., see the Open Source Initiative approved licenses: \href{https://opensource.org/licenses}{opensource.org/licenses}). This is an important step for improving reproducibility of research, «\emph{to provide the means by which the reader can verify the validity of the results and make use of them in further research}» \citep{GEO.17.Broggini}. Going a step beyond releasing open-source software, many groups adopt an open model of development where code is hosted and versioned in an online repository, all changes are public, and users can engage by submitting and track issues. Community oriented projects further engage by encouraging pull-requests, which are suggested changes to the code. Successful, well-maintained projects often have unit-testing and continuous integration that runs those tests with any changes to the code. Additionally, documentation that includes examples and tutorials is an important component for on-boarding new users and contributors. As a result, in many areas of the geosciences, we are seeing a shift away from a one-way distribution of code towards building global communities around open projects. Related projects within the realm of geophysical exploration are, e.g., \texttt{pyGIMLi} \citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}
% Others coming to mind?
% DW: - https://github.com/fancompute/ceviche
% RR: - no thoughts about geophysics, but there are many projects for
%       related topics such as gmt, gis, mesh-gen
% LH  - https://github.com/softwareunderground/awesome-open-geoscience ?
%     - Madagascar in seismic?

In this paper, we discuss the topic of open-source developments with respect to 3D CSEM modelling. Even though it comprises still the minority of developments, the landscape in this field of research changed in the last five years. Previously, there were only closed-source codes owned by companies or consortia and individual projects, often without proper documentation, that could be requested from the authors. We introduce and compare four recent open-source projects with focus on 3D CSEM modelling. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes.

@@

We simulate EM fields for a layered background model with vertical transverse isotropy, containing three blocks, as well as for the complex marine Marlim R3D model. These two model designs as well as the corresponding results from four different codes provide a benchmark for new codes to be compared to and validated with. In the following section, we introduce the codes under consideration. Afterwards, we present the considered benchmark cases in detail and present the modelling results of our four codes in terms of accuracy and computational performance. Beyond that, we extensively discuss important points that control the performance and suitability of our FV and FE codes, including mainly considerations about the mesh design and the choice of solvers. We conclude with a summary of our results and a motivation for the EM community to not only continue to extend the landscape of open-source codes but to also create a landscape of open-source benchmark models.

\itodo{Introduce Maxwell's Equations or not? Miensopust et al., 2013, don't.}
% Following the conventions of Ward and Hohmann \cite{Ward1988}, the quasi-static Maxwell's equations are

% \begin{equation}
% \begin{split}
%     \nabla \times \vec{E} &= -i \omega\vec{B}  \\
%     \nabla \times \vec{H} &= \vec{J}
% \end{split}
% \label{eq:maxwell_freq_quasi}
% \end{equation}

\section{Codes}

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg
\citep{CAG.15.Cockett, CAG.17.Heagy}. All four codes have their user-facing
routines written
in Python; all of them make heavy use of \texttt{NumPy}
\citep{CSE.11.VanDerWalt} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four
of them are “modern” open-source projects, meaning that they come with both, an
open-source license and an online-hosted version-control system with tracking
possibilities (raising issues, filing pull requests). All developments comprise
an extensive online documentation with many examples and have continuous
integration to some degree. Newer package-management systems such as Conda,
Docker, or pip significantly supported us to make our codes easily available
and installable without any user-side compilations. Each one can be downloaded
and installed with a single command.

In the following, we provide a quick overview of the codes. It is, however,
beyond the scope of this article to go into every detail of the different
modellers, and we refer to their documentations for more details. An overview
comparison of the different codes is given in Table~\ref{tbl:codecomparison}.
All codes have in common that they solve the weak formulation of Maxwell's
equation in its differential form (DE) under the quasistatic or diffusive
approximation, hence neglecting displacement currents. The machines on which
the different codes were run are listed in Table~\ref{tbl:machines} together
with the responsible operator.
% LH: SimPEG also interfaces to the scipy solvers; but in practice those are
%     use much less than Pardiso / mumps
% => I leave them out for now, as you did not use them for this paper either,
%    right? Maybe to mention in the SimPEG-section.

@@

\subsection{custEM}

The customisable electromagnetic modelling Python toolbox custEM was developed
for simulating arbitrary complex 3D CSEM geometries with focus on semi-airborne
setups, but it supports also land-based, airborne, coastal  or marine
environments. Multiple electric or magnetic field or potential finite-element
approaches were implemented as total or secondary field formulations. The
finite-element kernel, including higher order basis functions and
parallelisation, relies on the FEniCS project \citep{B.SPR.12.Logg,
B.SPR.16.Langtangen}. The resulting systems of linear equations are solved with
MUMPS \citep{SIAM.01.Amestoy}, which is a very robust but memory consuming
choice. Primary field solutions are supplied by the COMET \citep{GEO.20.Skibbe}
package.

The toolbox considers generally anisotropic petrophysical properties. Even
though changes of the conductivity is mainly of interest for CSEM modelling, the
electric permittivity and magnetic permeability can be taken into account using
the preferred electric field approach on Nédélec elements. Recently, induced
polarization parameters in frequency-domain and three methods for simulating
time-domain responses were added to custEM. The provided meshing tools based on
TetGen \citep{TOM.15.Si} and functionalities of pyGIMLi \citep{CAG.17.Rucker}
facilitate the generation of tetrahedral meshes including layered-earth
geometries with topography or bathymetry and anomalous bodies which are allowed
to be connected or to reach the surface.

The custEM toolbox is under steady development and maintenance. Future
enhancements comprise adding iterative solution techniques, provided by FEniCS,
as well as further functionalities for automatically incorporating intersecting
bodies such as folds or faults in meshes. Furthermore, current work is on using
custEM for inverse modelling applications.

@@

Based on the current state of the \petgem project, there are many possibilities
to improve or to add additional features. For instance, in the short-term, more
modelling routines and support for multi-source set-ups will be implemented. For
the long-term, the obvious next step is to move to 3D CSEM data inversion.

@@

\simpeg is a modular toolbox for simulations and gradient based inversions in
geophysics. Current functionality includes modelling and inversion
capabilities for gravity, magnetics, direct current resistivity, induced
polarization, electromagnetics (time, frequency, as well as natural sources,
e.g. magnetotellurics), and fluid flow (Richards equation). It is a community
driven project that aims to support researchers by providing a flexible,
extensible framework and to facilitate integration of data sets from a variety
of geophysical methods, including joint inversions \citep{GJI.20.Astic}, by
providing a common interface to each method.

Meshes and finite volume differential operators are implemented in the
\discretize package. It currently includes tensor, OcTree, cylindrical, and
logically rectangular meshes. Each mesh type inherits from a common structure
and uses the same naming conventions for methods and properties. This allows us
to decouple the implementation of a discretized partial differential equation
from the details of the mesh geometry and therefore we can write a single
implementation of discretized partial differential equation (or set of partial
differential equations) in \simpeg that will support all mesh types implemented
in \discretize.
